---
title: "Accessing and Analyzing Processed Data from R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(base)
library(rgdal)
library(MODIStsp)
library(dplyr)
```

## Accessing the processed time series from R

Preprocessed MODIS data can be retrieved within `R`  either by accessing the 
single-date raster files, or by loading the saved `RasterStack` objects (see 
[HERE](articles/output.html) for a description of `MODIStsp` output folder
structure and naming conventions).

To test this functionality, you can run the following example: 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, eval = FALSE}
library(MODIStsp)
opts_file <- system.file("testdata/test_extract.json", package = "MODIStsp")
MODIStsp(options_file = opts_file, gui = TRUE)
```

, specifying a folder on your PC to store the outputs from the GUI. 

This will download a yearly time series of MODIS NDVI data and subset it over the
region of the Como Lake in Italy.

After the download and processing finishes (it will take a while), the MODIS time 
series will be placed in `youroutfolder/VI_16Days_500m_v6/`. (In this case, I have them in 
"`D:/Temp/VI_16Days_500m_v6/`").

Any single-date image can be accessed by simply opening it with a `raster` command: 

```{R echo=TRUE, message=FALSE, warning=FALSE, highlight=TRUE, tidy=TRUE}
modistsp_file <- "D:/Temp/VI_16Days_500m_v6/NDVI/MOD13A1_NDVI_2016_177.tif"
my_raster     <- raster(modistsp_file)
plot(my_raster)
```

`RasterStack` (or GDAL vrt) time series containing all the processed data for a
given parameter are saved in the "Time Series/RData" subfolder, and can be accessed
by: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Load the NDVI time series
ts_folder <- "D:/Temp/VI_16Days_500m_v6/Time_Series/"
in_virtual_file <- file.path(ts_folder, "RData/Terra/NDVI/MOD13A1_NDVI_1_2016_353_2016_RData.RData")
ts_data          <- get(load(in_virtual_file))
ts_data
```

, which gives us a 23-band `RasterStack` in the `ts_data` variable.

This `RasterStack` can be analyzed using the functionalities for raster/raster 
time series analysis, extraction and plotting provided for example by the 
```raster``` or   ```rasterVis``` packages: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# plot some dates
plot(ts_data[[c(1,5,10,15,20)]])
```

________________________________________________________________________________

## Extracting Time Series Data on Areas of Interest

`MODIStsp` provides an efficient function (`MODIStsp_extract`) for extracting 
time series data at specific locations. The function takes as input a `RasterStack` 
virtual object created by `MODIStsp` (see above), the starting and ending dates 
for the extraction and a standard `sp` object (or an ESRI shapefile name) 
specifying the locations (points, lines or polygons) of interest, and provides 
as output a `xts` object or `data.frame` containing time series data for those
locations. 

If the input is of class _SpatialPoints_, the output object contains one column 
for each point specified, and one row for each date. If it is of class
_SpatialPolygons_ (or _SpatialLines_), it contains one column for each polygon 
(or each line), with values obtained applying the function specified as the "FUN" 
argument (e.g., mean, standard deviation, etc.) on pixels belonging to the polygon
(or touched by the line), and one row for each date. 

To test `MODIStsp_extract` we can start by loading the NDVI `RasterStack`
time series created in the previous example:

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

ts_folder <- "D:/Temp/VI_16Days_500m_v6/Time_Series/"
in_virtual_file <- file.path(ts_folder, "RData/Terra/NDVI/MOD13A1_NDVI_1_2016_353_2016_RData.RData")
ts_data          <- get(load(in_virtual_file))
ts_data

```

, which again gives us the 23-band `RasterStack` in the `ts_data` variable. 

To extract the NDVI data over selected areas, we can then use the `MODIStsp_extract`
function as follows: 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# load a shapefile containing polygons --> Here we use a test shapefile in 
# the testdata folder of MODIStsp.

polygons <- rgdal::readOGR(system.file("testdata/extract_polys.shp",
                                       package = "MODIStsp"),
                           verbose = FALSE)
polygons

# Now extract the average values for each polygon and date and plot the results

out_dataavg <- MODIStsp_extract(ts_data, polygons, id_field = "lc_type", 
                                small = FALSE)
head(out_dataavg)

# Other summarization functions can be used, by specifying the "FUN" argument. 
# Compute the Standard Deviation over each polygon: 
out_datasd <- MODIStsp_extract(ts_data, polygons, id_field = "lc_type",
FUN = "sd", small = FALSE)

```

The output is a `xts` object, with one column for each polygon of the input 
shapefile, and one row for each date. We can easily plot the computed averages 
using: 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(xts)
plot(out_dataavg, legend.loc = "topleft")
```

We can also transform the output dataset to a long format and use `ggplot2` for
plotting and other tidyverse-friendly functions for analysis:

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
out_dataavg_df <- data.frame(date = index(out_dataavg), coredata(out_dataavg)) %>% 
  tidyr::gather(Polygon, Value, -date)

ggplot2::ggplot(out_dataavg_df, aes(x = date, y = Value, color = Polygon)) + 
  geom_line() + theme_light()

```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(tibbletime)
library(dplyr)

# Compute monthly averages using tibbletime: 
 
out_dataavg_tt <- tibbletime::as_tbl_time(out_dataavg_df, index = date) 

month_avg <- out_dataavg_tt %>% 
  dplyr::group_by(Polygon) %>% 
  tibbletime::as_period(period = "monthly")

ggplot2::ggplot(month_avg, aes(x = date, y = Value, color = Polygon)) + 
  geom_line() + theme_light()

```

**IMPORTANT NOTES - 1** `MODIStsp_extract` is usually much faster than the 
standard `raster::extract` function, but does not deal with overlapping polygons.
If your polygons are overlapping, please use `raster::extract` instead. 

**IMPORTANT NOTES - 2** A more powerful extraction function can be also found 
in the `sprawl` package, currently under development on github
(https://lbusett.github.io/sprawl/). In particular, `sprawl::extract_rast` deals
much better with large rasters and long time series. 

