---
title: "Accessing and Analyzing Processed Data from R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(base)
```

## Accessing the processed time series from R

Preprocessed MODIS data can be retrieved within `R`  either by accessing the single-date raster files, or by loading the saved _RasterStack_ objects. 

Any single-date image can be accessed by simply opening it with a `raster` command: 

```{R eval=FALSE, tidy=TRUE, highlight=TRUE}
library(raster)
modistsp_file <- "/my_outfolder/EVI/MOD13Q1_2005_137_EVI.tif"
my_raster     <- raster(modistsp_file)
```

`rasterStack` time series containing all the processed data for a given parameter (saved in the "Time Series/RData" subfolder - see [here](output.html) for details) can be opened by: 

```{r eval=FALSE}
in_virtual_file <- "/my_outfolder/Time_Series/RData/Terra/EVI/MOD13Q1_MYD13Q1_NDVI_49_2000_353_2015_RData.RData" 
indata          <- get(load(in_virtual_file))
```

This second option allows accessing the complete data stack and analyzing it using the functionalities for raster/raster time series analysis, extraction and plotting provided for example by the ```raster``` or   ```rasterVis``` packages.

____________________________________________________________________________________

## Extracting Time Series Data on Areas of Interest

`MODIStsp` provides an efficient function (`MODIStsp\_extract`) for extracting time series data at specific locations. The function takes as input a _RasterStack_ virtual object created by `MODIStsp` (see above), the starting and ending dates for the extraction and a standard _Sp*_ object (or an ESRI shapefile name) specifying the locations (points, lines or polygons) of interest, and provides as output a `xts` object or `data.frame` containing time series data for those locations. 

If the input is of class _SpatialPoints_, the output object contains one column for each point specified, and one row for each date. If it is of class _SpatialPolygons_ (or _SpatialLines_), it contains one column for each polygon (or each line), with values obtained applying the function specified as the "FUN" argument (e.g., mean, standard deviation, etc.) on pixels belonging to the polygon (or touched by the line), and one row for each date. 

As an example the following code:

```{r, eval=FALSE}
  #Set the input paths to raster and shape file
  infile    <- 'myoutfolder/Time_Series/RData/Mixed/MOD13Q1_MYD13Q1_NDVI_49_2000_353_2015_RData.RData'  
  shpname   <- 'path_to_file/rois.shp'  
  #Set the start/end dates for extraction
  startdate <- as.Date("2010-01-01")  
  enddate   <- as.Date("2014-12-31")    
  #Load the RasterStack
  inrts     <- get(load(infile)) 
  # Compute average and St.dev
  dataavg   <- MODIStsp_extract(inrts, shpname, startdate, enddate, FUN = 'mean', na.rm = T)
  datasd    <- MODIStsp_extract (inrts, shpname, startdate, enddate, FUN = 'sd', na.rm = T)
  # Plot average time series for the polygons
  plot.xts(dataavg) 
```

  loads a _RasterStack_ object containing 8-days 250 m resolution time series for the 2000-2015 period and extracts time series of average and standard deviation values over the different polygons of a user's selected shapefile on the 2010-2014 period.
  