---
title: "Accessing and Analyzing Processed Data from R"
output: 
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(knitr)
library(raster)
library(base)
library(sf)
library(MODIStsp)
library(dplyr)
test_zip <-  system.file("testdata/VI_16Days_500m_v61/NDVI.zip",
                         package = "MODIStsp")
dir.create(file.path(tempdir(), "MODIStsp/VI_16Days_500m_v61"),
           showWarnings = FALSE, recursive = TRUE)
utils::unzip(test_zip, 
             exdir = file.path(tempdir(), "MODIStsp/VI_16Days_500m_v61"))
Sys.setlocale("LC_TIME", "en_US.UTF-8")
```

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Accessing the raster time series__

Preprocessed MODIS data can be retrieved within `R` either by accessing the 
single-date raster files, or by loading the saved `RasterStack` objects (see 
[here](output.html) for a description of `{MODIStsp}` output folder
structure and naming conventions).

To test this functionality, you can run the following example: 

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(MODIStsp)
opts_file <- system.file("testdata/test_extract.json", package = "MODIStsp")
MODIStsp(opts_file = opts_file, gui = FALSE, verbose = FALSE)
```

This will download a yearly time series of MODIS NDVI data and subset it over the
region of the Como Lake in Italy.

After the download and processing finishes (it will take a while, depending on your
network speed), the MODIS time series will be placed in subfolder `MODIStsp/VI_16Days_500m_v61` 
of `R` `tempdir()`. In this case, I have them in "`file.path(tempdir(), "VI_16Days_500m_v61")`". 
If you want to save them elsewhere, run `MODIStsp(opts_file = opts_file, gui = TRUE)`
and select a different output folder.

Single-date processed rasters can be accessed by simply opening them with a
`raster` command: 

```{R echo=TRUE, message=FALSE, warning=FALSE, highlight=TRUE, tidy=TRUE}
modistsp_file <- file.path(tempdir(),"MODIStsp/VI_16Days_500m_v61/NDVI",
                           "MOD13A1_NDVI_2016_177.tif")
modistsp_file

my_raster     <- raster(modistsp_file)
plot(my_raster)
```

`RasterStack` (or GDAL vrt) time series containing all the processed data for a
given parameter are saved in the `Time Series/RData` subfolder, and can be accessed
by: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Load the NDVI time series
ts_folder <- file.path(tempdir(), "MODIStsp/VI_16Days_500m_v61/Time_Series")
in_virtual_file <- file.path(ts_folder, "RData/Terra/NDVI/MOD13A1_NDVI_1_2016_353_2016_RData.RData")
in_virtual_file

ts_data   <- get(load(in_virtual_file))
ts_data
```

, which gives us a 23-band `RasterStack` in the `ts_data` variable.

This `RasterStack` can be analyzed using the functionalities for raster/raster 
time series analysis, extraction and plotting provided for example by the 
``{raster}`` or   ``{rasterVis}`` packages: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# plot some dates
plot(ts_data[[c(1,5,10,15,20)]], axes = FALSE, horizontal = T)

# Extract one date from the stack

mydate   <- as.Date("2016-01-01")
substack <- subset(ts_data, which(getZ(ts_data) == mydate)) %>% setZ(mydate)
substack  

# Extract multiple dates from the stack

mindate   <- as.Date("2016-01-01")
maxdate   <- as.Date("2016-04-01")
substack  <- subset(ts_data, 
                    which(getZ(ts_data) >= mindate & getZ(ts_data) <= maxdate))
substack  

# Compute monthly averages

month_avg <- stackApply(ts_data, months(getZ(ts_data)), fun = mean)
month_avg

plot(month_avg, axes = FALSE, horizontal = T)
```

________________________________________________________________________________

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Extracting Time Series Data on Areas of Interest__

`{MODIStsp}` provides an efficient function (`MODIStsp_extract()`) for extracting 
time series data at specific locations. The function takes as input a `RasterStack` 
virtual object created by `MODIStsp()` (see above), the starting and ending dates 
for the extraction and a standard `{sp}` object (or an ESRI shapefile name) 
specifying the locations (points, lines or polygons) of interest, and provides 
as output a `xts` object or `data.frame` containing time series data for those
locations. 

If the input is of class `SpatialPoints`, the output object contains one column 
for each point specified, and one row for each date. If it is of class
`SpatialPolygons` (or `SpatialLines`), it contains one column for each polygon 
(or each line), with values obtained applying the function specified as the "FUN" 
argument (e.g., mean, standard deviation, etc.) on pixels belonging to the polygon
(or touched by the line), and one row for each date. 

To test `MODIStsp_extract()` we can start by loading the NDVI `RasterStack`
time series created in the previous example:

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

ts_folder <- file.path(tempdir(), "MODIStsp/VI_16Days_500m_v61/Time_Series")
in_virtual_file <- file.path(ts_folder, "RData/Terra/NDVI/MOD13A1_NDVI_1_2016_353_2016_RData.RData")
ts_data   <- get(load(in_virtual_file))
ts_data
```

, which again gives us the 23-band `RasterStack` in the `ts_data` variable. 

To extract the NDVI data over selected areas, we can then use the `MODIStsp_extract()`
function as follows: 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# load a shapefile containing polygons --> Here we use a test shapefile in 
# the testdata folder of MODIStsp.

polygons <- sf::st_read(system.file("testdata/extract_polys.shp",
                                       package = "MODIStsp"),
                           quiet = TRUE)
polygons

# Now extract the average values for each polygon and date and plot the results

out_dataavg <- MODIStsp_extract(ts_data, polygons, id_field = "lc_type", 
                                small = FALSE)

head(out_dataavg)

# Other summarization functions can be used, by specifying the "FUN" argument. 
# Compute the Standard Deviation over each polygon: 
out_datasd <- MODIStsp_extract(ts_data, polygons, id_field = "lc_type",
                               FUN = "sd", small = FALSE)

```

The output is a `xts` object, with one column for each polygon of the input 
shapefile, and one row for each date. We can easily plot the computed averages 
using: 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(xts)
plot(out_dataavg, legend.loc = "topleft")
```

We can also transform the output dataset to a long format and use `{ggplot2}` for
plotting and other tidyverse-friendly functions for analysis:

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
out_dataavg_df <- data.frame(date = index(out_dataavg), coredata(out_dataavg)) %>% 
  tidyr::gather(Polygon, Value, -date)

ggplot2::ggplot(out_dataavg_df, aes(x = date, y = Value, color = Polygon)) + 
  geom_line() + theme_light()

```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

require(tibbletime)
require(dplyr)

# Compute monthly averages using tibbletime: 

out_dataavg_tt <- tibbletime::as_tbl_time(out_dataavg_df, index = date) 

month_avg <- out_dataavg_tt %>% 
  dplyr::group_by(Polygon) %>% 
  tibbletime::as_period(period = "monthly")

ggplot2::ggplot(month_avg, aes(x = date, y = Value, color = Polygon)) + 
  geom_line() + theme_light()

```

**Important note:** `MODIStsp_extract()` is usually much faster than the 
standard `raster::extract()` function, but does not deal with overlapping polygons.
If your polygons are overlapping, please use `raster::extract()` instead. 
