---
title: Installation
output: 
  github_document:
    toc_dept: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <i class="fa fa-desktop" aria-hidden="true"></i> System Requirements 

`MODIStsp` requires [R](https://cran.r-project.org) v >= 3.2.1 and [GDAL](http://www.gdal.org) (Geospatial Data Abstraction Library) v >= 1.11.1 __with support for HDF4 raster format__ to be installed in your system. Brief instructions for installing R and GDAL can be found [HERE](#installing-r-and-gdal).

____________________________________________________________________________________

## <i class="fa fa-windows" aria-hidden="true"></i> Installing on Windows

You can install the stable version of `MODIStsp`, from CRAN: 

`install.packages("MODIStsp")`

, or the development version (containing the latest improvements and bug fixes):

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("lbusett/MODIStsp")
```

Note that **if the `GTK+` library is not already installed on your system, installation may fail**. In that case, please install and load the `gWidgetsRGtk2` library beforehand:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("gWidgetsRGtk2")
library(gWidgetsRGtk2)
```
    
 Upon loading `gWidgetsRGtk2` , an error window will probably appear. This signals that library "GTK+" is not yet installed on your system or is not on your PATH. To install it press "OK". A new window dialog window will appear, asking if you want to install "GTK+". Select "Install GTK" and then "OK" . Windows will download and install the GTK+ library. When it finishes, the RSession should be restarted and you should be ready to go !<sup name="a1">
 
 In case RStudio doesn't automatically restart or continuously asks to install GTK+ again, kill it form "Task Manager" (or restart the R session from RStudio "Session" menu), reload RStudio and the try to reload `gWidgetsRGtk2`. If it loads correctly, you should be ready to go. 

If it still fails, try downloading the GTK+ bundle from:

<http://ftp.gnome.org/pub/gnome/binaries/win64/gtk+/2.22/gtk+-bundle_2.22.1-20101229_win64.zip> 
(OR <http://ftp.gnome.org/pub/gnome/binaries/win32/gtk+/2.22/gtk+-bundle_2.22.1-20101227_win32.zip> if on Win32)

, unzip the archive on a folder of your choice (e.g., `C:\\Program Files\\GTK+`), then add the path to its "bin" subfolder (e.g., `C:\\Program Files\\GTK+\\bin\\` to your system PATH environment variable.

Restart your system and try loading again `gWidgetsRGtk2`: if it loads ok, you should be ready to install `MODIStsp`

____________________________________________________________________________________

## <i class="fa fa-linux" aria-hidden="true"></i> Installing on Linux Systems

To install `MODIStsp` on Linux, you have to first install the following required dependencies: 

  * `Cairo` >= 1.0.0, `ATK` >= 1.10.0, `Pango` >= 1.10.0, `GTK+` >= 2.8.0, `GLib` >= 2.8.0 (required by package ```RGtk2```)
  * `Curl` (required by package ```curl```)
  * `GDAL` >= 1.6.3, `PROJ.4` >= 4.4.9 (required by package ```rgdal```)
    
On *Debian and Ubuntu-based* systems, to install those packages open a terminal and type: 
    
```bash
sudo apt-get install r-cran-cairodevice r-cran-rgtk2 libcairo2-dev libatk1.0-dev libpango1.0-dev 
libgtk2.0-dev libglib2.0-dev libcurl4-openssl-dev libgdal-dev libproj-dev
```
    
On *rpm-base systems*, to install packages open a terminal and type:  

```bash
sudo yum install libcairo2-devel libatk1.0-devel libpango1.0-devel gtk2 gtk2-devel 
glib2-devel libcurl-devel gdal-devel proj proj-devel proj-epsg proj-nad
```
    
Then, you can install the stable version of MODIStsp from CRAN:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("MODIStsp")
```
, or the development version (containing the latest improvements and bug fixes) from github;

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("lbusett/MODIStsp")
```
____________________________________________________________________________________

## <i class="fa fa-apple" aria-hidden="true"></i> Installing on Mac

**NOTE: The following  installation notes were kindly provided by a MODIStsp user
and should be valid for MODIStsp installation on R 3.4.0 and above with 
Mac OSX Sierra. Since we are not working on Mac we were not able yet to
check this, but we hope it may be useful at least to put you in the right 
direction if you have installation problems)**

Installing MODIStsp requires many dependencies: 

For installation on MAC OSX sierra, there are three main issues:
- As outlined here in the comment by **tobybot11** (https://gist.github.com/sebkopf/9405675), 
Rgtk requires the x11 libraries/headers (gdk/gdkx.h specifically) and doesn't
work with the quartz libraries/headers which now are the default for GTK
- When installing the dependencies gWidgetsRGtk2 and cairoDevice from CRAN you
need to choose the version "OS X Mavericks binaries" and not "Package source"
- You have to be sure that gdal is installed with HDF4 support.

1 - Update to R > 3.4  if needed, then update all packages

```{r eval = FALSE}
# update packages
update.packages()
# install the development version of devtools:
install.packages(c("devtools"))
devtools::install_github("hadley/devtools")
```

2 - Now, install RGtk2 using Homebrew (https://gist.github.com/sebkopf/9405675).
First, ensure you have cairo installed with "--with-x11". Open a terminal and 
run:

```bash
brew uninstall cairo --ignore-dependencies
brew install --with-x11 cairo 
```
next, edit the configure options for GTK to require x11 rather than Quartz:
```bash
brew edit gtk+
```
in the def install section, remove the reference to quartz and switch to: 
```
--with-gdktarget=x11,
--enable-x11-backend
```
Now install: 
```bash
brew install --build-from-source --verbose gtk+
export PKG_CONFIG_PATH=/usr/local/lib/pkgconfig:/usr/local/lib/pkgconfig/gtk+-2.0.pc:/opt/X11/lib/pkgconfig
```

3 - next, install `RGtk2`. Since install.packages("Rgtk2") is not going to work, go [here](https://cran.r-project.org/web/packages/RGtk2/index.html), download 
RGtk2_2.20.33.tar.gz and, from a terminal run:
```bash
R CMD INSTALL RGtk2_2.20.33.tar.gz
```
Now, open R and run: 

```{r eval = FALSE}
library(RGtk2)
```
hopefully, `RGTk2` will load without errors !

4 - Install packages `gWidgetsRGtk2` and `cairoDevice`

**Very important !!!!** from CRAN, download the **"OS X Mavericks binaries"** for 
[gWidgetsRGtk2](https://cran.r-project.org/web/packages/gWidgetsRGtk2/index.html) and [cairoDevice](https://cran.r-project.org/web/packages/cairoDevice/index.html)(
not "Package source"). Save both to Frameworks/R.framework/3.4..., open R and
run the code below (This will also install cairoDevice) 

```{r eval = FALSE}
install.packages("gWidgetsRGtk2",
                 lib="~/Library/Frameworks/R.framework/Versions/3.4/Resources/library/gWidgetsRGtk2")
library(gWidgetsRGtk2)
library(cairoDevice)
```
_(This will work for R version 3.4, mac OS X Sierra)_

5 - Install GDAL with HDF4 support

Check that gdal is installed with hdf4 support. From a terminal: 
```bash
> gdal-config --formats 
```
if gdal is installed, check what drivers are installed: the list should include
hdf4. 

If gdal is not yet installed or hdf4 is not supproted, install/reinstall it following 
these [notes](http://r-sig-geo.2731867.n2.nabble.com/OSX-rgdal-GDAL-drivers-and-HDF4-td7588400.html)
```bash
> brew install hdf4 
# prefer hdf4 links over NetCDF 
> brew link --overwrite hdf4 
> brew install gdal --complete --enable-unsupported --with-hdf4 
# check what drivers are installed, list should now include hdf4: 
> gdal-config --formats 
```
6 - Install rgdal

since rgdal is not usually looking in "/usr/local/lib"" you must include that with
`install.packages()`: 

```{r eval = FALSE}
install.packages('rgdal',
                 type = "source",configure.args = c(
                   '--with-proj-include=/usr/local/include',
                   '--with-proj-lib=/usr/local/lib')
                 )
```
7 - Finally, install MODIStsp from CRAN or GitHub:

```{r eval = FALSE}
install.packages("MODIStsp")
MODIStsp()
```
OR
```{r eval = FALSE}
library(devtools) 
install_github("lbusett/MODIStsp", ref = "master")
MODIStsp()
```
Good luck!
________________________________________________________________________________

# Installing R and GDAL

### Installing R

#### Windows

Download and install the latest version of R which can be found [here](https://cran.r-project.org/bin/windows/base).

#### Linux

Please refer to the documentation which can be found [here](https://cran.r-project.org/bin/linux), opening the directory relative to your Linux distribution. The documentation provides instruction to add CRAN repositories and to install the latest R version.
With Ubuntu 15.10 Wily (and newer) this step is not mandatory (although recommended), since packaged version of R is >= 3.2.1 (although not the latest); in this case, user can install R by simply typing in a terminal
```bash
sudo apt-get install r-base
```

### Installing GDAL >= 1.11.1

#### Windows

The easiest way to install GDAL on Windows is from the [OSGeo4W Website](https://trac.osgeo.org/osgeo4w/)

1. Open the [OSGeo4W Website](https://trac.osgeo.org/osgeo4w/)
2. In the **Quick Start for OSGeo4W Users** section, select the download of 32bit or 64bit of OSGeo4W network installer
3. Run the installer
  + _Easiest Option_: 
    +  Select **Express Desktop Install**, then proceed with the installation. This will install GDAL and also other useful Spatial Processing software like [QGIS](http://www.qgis.org/) and [GRASS GIS](https://grass.osgeo.org/)
  + _Advanced Option_: 
    +  Select **Advanced Install**, then click on "Next" a few times until you reach the "Select Packages" screen. 
    +  Click on "Commandline_Utilities_", and on the list look for "_gdal: The GDAL/OGR library..." entry
    +  Click on "Skip": the word "skip" will be replaced by the current GDAL version number
    +  Click on "Next" a few times to install GDAL
    
#### Debian and Ubuntu-based systems

1. Ensure that your repositories contain a version of ```gdal-bin``` >= 1.11.1. 
    In particular, official repositories of Ubuntu 15.04 Vivid (or older) and Debian Jessie (or older) provide older versions of GDAL, so it is necessary to add UbuntuGIS-unstable repository before installing. To do this, follow instructions [here](https://launchpad.net/~ubuntugis/+archive/ubuntu/ubuntugis-unstable)). 
    With Ubuntu 15.10 Wily (and newer) this step is not mandatory, although recommended in order to have updated version of GDAL installed.
    
2. To install GDAL, open a terminal and type  
    ```bash
    sudo apt-get install gdal-bin
    ```
    
#### ArchLinux

GDAL is maintained updated to the latest version as binary package within the community repository; although that, the support for HDF4 format is not included. 
To bypass this problem, ArchLinux users can install ```gdal-hdf4``` package from AUR (see [here](https://wiki.archlinux.org/index.php/Arch_User_Repository#Installing_packages) or [here](https://archlinux.fr/yaourt-en) for the package installation from AUR). 
This package is updated manually after each release of ```gdal``` on the community repository, so a temporal shift between a new ```gdal``` release and the update of ```gdal-hdf4``` could happen.
If you want to manually add the support for HDF4 in case ```gdal-hdf4``` is out-of-date, you can do it following [these instructions](https://notehub.org/fctdn).

#### Other Linux systems

Install the packaged binary of GDAL included in your specific distribution; if the version is older than 1.11.1, or if the support for HDF4 format is not included, you can manually install the HDF4 library and compile the source code by adding the parameter ```--with-hdf4``` to the ```configure``` instruction).

#### MacOs