---
title: Installation
output: 
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <i class="fa fa-windows" aria-hidden="true"></i> Installing on Windows

You can install the stable version of `{MODIStsp}` from CRAN: 

`install.packages("MODIStsp")`

, or the development version (containing the latest improvements and bug fixes) 
from GitHub:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("remotes")
library(remotes)
install_github("ropensci/MODIStsp")
```

## <i class="fa fa-linux" aria-hidden="true"></i> Installing on Linux Systems

To install `{MODIStsp}` on Linux, you need to be able to install the `{sf}` package, 
which requires several dependencies. See [HERE](https://github.com/r-spatial/sf#installing)
if you have trouble installing `{sf}`.

In addition, you need to install dependencies
required by the `{protolite}` package, required by `{geojson}`. See [HERE](https://github.com/jeroen/protolite/) for instructions
on installing them. 

Then, you can install the stable version of `{MODIStsp}` from CRAN:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("MODIStsp")
```
, or the development version (containing the latest improvements and bug fixes) 
from GitHub:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("ropensci/MODIStsp")
```

Notice that, in order to be able to read input HDF4 MODIS archives, your GDAL
installation must support the HDF4 format.
You can check that with the command:
```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
"HDF4" %in% sf::st_drivers("raster")$name
```
, that should return TRUE.
Most of the OS includes a GDAL distribution which supports this format, 
while some do not
(e.g. ArchLinux, see [here](https://github.com/ropensci/MODIStsp/issues/161)).

## <i class="fa fa-apple" aria-hidden="true"></i> Installing on Mac

To install `{MODIStsp}` on MacOS, you need to be able to install the `{sf}` package, 
which requires gdal to be installed. See [HERE](https://github.com/r-spatial/sf#installing)
if you have trouble installing `{sf}`. 

Then, you can install the stable version of `{MODIStsp}` from CRAN:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("MODIStsp")
```
, or the development version (containing the latest improvements and bug fixes) 
from GitHub:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("ropensci/MODIStsp")
```
