---
title: 'MODIStsp: A Tool for Automatic Preprocessing of MODIS Time Series - v1.3.6'
author: "Lorenzo Busetto ([busetto.l@irea.cnr.it](mailto:lbusett@gmail.com)), Luigi
  Ranghetti ([ranghetti.l@irea.cnr.it](mailto:ranghetti.l@irea.cnr.it))"
bibliography: MODIStsp.bib
date: '`r Sys.Date()`'
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  rmarkdown::html_document:
    fig_caption: yes
    number_section: yes
    toc: no
    toc_depth: 1
  rmarkdown::html_vignette:
    fig_caption: yes
    number_section: yes
    toc: no
    toc_depth: 2
urlcolor: blue
linkcolor: blue
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{"MODIStsp: A Tool for Automatic Preprocessing of MODIS Time Series - v1.3.6"}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

MODIStsp is a novel "R" package allowing to automatize the creation of time series
of rasters derived from MODIS Land Products data. It allows to perform several 
preprocessing steps on MODIS data available within a given time period. 

Development of MODIStsp started from modifications of the ModisDownload "R" script
by Thomas Hengl [-@Hengl2010], and successive adaptations by Babak Naimi [-@Naimi2014]. 
The basic functionalities for download and preprocessing of MODIS datasets provided 
by these scripts were gradually incremented with the aim of: 

* developing a stand-alone application allowing to perform several preprocessing
steps (e.g., download, mosaicing, reprojection and resize) on all available MODIS 
land products by exploiting  a powerful and user-friendly GUI front-end;
* allowing the creation of time series of both MODIS original layers and additional 
Quality Indicators (e.g., data acquisition quality, cloud/snow presence, algorithm 
used for data production, etc. ) extracted from the aggregated bit-field QA layers;
* allowing the automatic calculation and creation of time series of several additional
Spectral Indexes starting form MODIS surface reflectance products.

All processing parameters can be easily set with a user-friendly GUI, although
non-interactive execution exploiting a previously created Options File is possible. 
Stand-alone execution outside an "R" environment is also possible, allowing to use
scheduled execution of MODIStsp to automatically update time series related to a 
MODIS product and extent whenever a new image is available. 

Required MODIS HDF files are automatically downloaded from NASA servers and resized,
reprojected, resampled and processed according to user's choices. For each desired 
output layer, outputs are saved as single-band rasters corresponding to each acquisition 
date available for the selected MODIS product within the specified time period. 
"R" _RasterStack_ objects with temporal information as well as Virtual raster files
(GDAL vrt and ENVI META files) facilitating access to the entire time series can
be also created.

# Installation

`MODIStsp` requires [R](https://cran.r-project.org) v >= 3.2.1 and [GDAL](http://www.gdal.org) (Geospatial Data Abstraction Library) v >= 1.11.1 __with support for HDF4 raster format__ to be
installed in your system. Brief instructions for installing R and GDAL can be found [HERE](#installing-r-and-gdal).

## On Windows

You can install the stable version of `MODIStsp`, from CRAN: 

`install.packages("MODIStsp")`

, or the development version (containing the latest improvements and bug fixes):

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("lbusett/MODIStsp")
```

Note that **if the `GTK+` library is not already installed on your system, installation may fail**.
In that case, please install and load the `gWidgetsRGtk2` library beforehand:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("gWidgetsRGtk2")
library(gWidgetsRGtk2)
```
    
 Upon loading `gWidgetsRGtk2` , an error window will probably appear. This signals
 that library "GTK+" is not yet installed on your system or is not on your PATH.
 To install it press "OK". A new window dialog window will appear, asking if you 
 want to install "GTK+". Select "Install GTK" and then "OK" . Windows will download
 and install the GTK+ library. When it finishes, the RSession should be restarted
 and you should be ready to go !<sup name="a1">
 
 In case RStudio does not automatically restart or continuously asks to install
 GTK+ again, kill it form "Task Manager" (or restart the R session from RStudio
 "Session" menu), reload RStudio and the try to reload `gWidgetsRGtk2`. If it
 loads correctly, you should be ready to go. 

If it still fails, try downloading the GTK+ bundle from:

<http://ftp.gnome.org/pub/gnome/binaries/win64/gtk+/2.22/gtk+-bundle_2.22.1-20101229_win64.zip> 
(OR <http://ftp.gnome.org/pub/gnome/binaries/win32/gtk+/2.22/gtk+-bundle_2.22.1-20101227_win32.zip> if on Win32)

, unzip the archive on a folder of your choice (e.g., `C:\\Program Files\\GTK+`), then add the path to its "bin" subfolder (e.g., `C:\\Program Files\\GTK+\\bin\\` to your system PATH environment variable.

Restart your system and try loading again `gWidgetsRGtk2`: if it loads ok, you should be ready to install `MODIStsp`

    
## On Linux systems

To install `MODIStsp` on Linux, you have to first install the following required dependencies: 

  * `Cairo` >= 1.0.0, `ATK` >= 1.10.0, `Pango` >= 1.10.0, `GTK+` >= 2.8.0, `GLib` >= 2.8.0 (required by package ```RGtk2```)
  * `Curl` (required by package ```curl```)
  * `GDAL` >= 1.6.3, `PROJ.4` >= 4.4.9 (required by package ```rgdal```)
    
On *Debian and Ubuntu-based* systems, to install those packages open a terminal and type: 
    
```bash
sudo apt-get install r-cran-cairodevice r-cran-rgtk2 libcairo2-dev libatk1.0-dev libpango1.0-dev 
libgtk2.0-dev libglib2.0-dev libcurl4-openssl-dev libgdal-dev libproj-dev
```
    
On *rpm-base systems*, to install packages open a terminal and type:  

```bash
sudo yum install libcairo2-devel libatk1.0-devel libpango1.0-devel gtk2 gtk2-devel 
glib2-devel libcurl-devel gdal-devel proj proj-devel proj-epsg proj-nad
```
    
Then, you can install the stable version of MODIStsp from CRAN:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("MODIStsp")
```
, or the development version (containing the latest improvements and bug fixes) from github;

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("lbusett/MODIStsp")
```
## On Mac OS

**NOTE**: The following  installation notes should be valid for MODIStsp installation 
on R 3.4.0 and above with Mac OSX Sierra. They were mainly taken (i.e., blatantly copied...) from: https://zhiyzuo.github.io/installation-rattle/. Thanks to Zhiya Zuo for providing this!

To properly install `MODIStsp` you will need to first install package `RGTk2`. This
is a somehow difficult operation. The following instructions should help:

__1. Check your Mac OS X version and update if necessary: __

Enter the following command in terminal to check your macOS version. Expected
output is as below the dashed line ---.
    
```
~$ sw_vers  
------------------------  
ProductName:	Mac OS X  
ProductVersion:	10.12.6  
BuildVersion:	16G29  
```

If your system is above 10.11, continue. Otherwise, upgrade it to Sierra.

Install homebrew if you do not have it already installed. homebrew  is a very
convenient package manager for macOS. To do so, open a terminal, copy the following
command in it and hit Enter:

```
~$ /usr/bin/ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)"
```

Follow the instructions to get brew ready. When inserting your password, nothing 
will show up for security reasons. Just hit Enter when you are finished.

When brew is finished, copy the following command in terminal and hit Enter:

```
~$ touch ~/.bash_profile
~$ echo "export PATH=/usr/local/bin:$PATH
export PKG_CONFIG_PATH=/usr/local/lib/pkgconfig:/usr/local/lib/pkgconfig/gtk+-2.0.pc:/opt/X11/lib/pkgconfig" >> ~/.bash_profile
~$ source ~/.bash_profile
```


__2. Install the `cairo` library with x11 support__. Enter the following into your terminal:
 
```
~$ brew uninstall --force cairo --ignore-dependencies
~$ brew cask install xquartz
~$ brew install --with-x11 cairo
```

 
__3. Install the `gtk+` library__:
 
To do so, you first have to change the way homebrew wants to install gtk+. In an editor, write: 
 
```
~$ brew edit gtk+
```
 
A text editor will open. Look in the file, and find a section that begins with
"def install". Substitute the current `args` section with the following text: 
 
```
def install
 args = [
         "--disable-dependency-tracking",
         "--disable-silent-rules",
         "--prefix=#{prefix}",
         "--disable-glibtest",
         "--enable-introspection=yes",
         # "--disable-visibility",
         # "--with-gdktarget=quartz",
         "--with-gdktarget=x11",
         "--enable-x11-backend"
        ]
```     
 
Save the modified file using `ctrl+x ctrl+c`, followed by `y` to quit emacs.
Now install the library using: 
 
```
~$ brew install --build-from-source --verbose gtk+
```


__4. Update your path__ so that `gtk+` is recognized, using:

```
~$ export PKG_CONFIG_PATH=/usr/local/lib/pkgconfig:/usr/local/lib/pkgconfig/gtk+-2.0.pc:/opt/X11/lib/pkgconfig
```


5. __Install `RGtk2` from source__:

- __Download the newest source file for RGtk2__ from https://CRAN.R-project.org/package=gWidgetsRGtk2.

- Assuming that the path to this file is ~/Downloads. Run the following in terminal 
(change the path if you did not download in ~/Downloads):

```
~$ cd ~/Downloads
~/Downloads$ R CMD INSTALL RGtk2_2.20.33.tar.gz
```
(Note that the name of the tar.gz file may vary depending on when you downloaded the file). 


6. __Open R and run__: 

```{r eval = FALSE}
library(RGtk2)
```
hopefully, `RGtk2` will load without errors! If so, you should be ready to go, 
and you can:

7. __Install MODIStsp__ from CRAN:

```{r eval = FALSE}
install.packages("MODIStsp")
MODIStsp()
```
or the development version from GitHub:
```{r eval = FALSE}
library(devtools) 
install_github("lbusett/MODIStsp", ref = "master")
MODIStsp()
```
Good luck!

# Running the tool in Interactive Mode: the MODIStsp GUI

The easiest way to use `MODIStsp` is to use its powerful GUI (Graphical User 
Interface) for selection of processing options, and then run the processing. 

To open the GUI, load the package and launch the MODIStsp function, with no parameters:
```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, caption=FALSE}
library(MODIStsp)
MODIStsp()
```
This **opens a GUI** from which processing options can be specified and eventually
saved (or loaded) 
<font size="2"> (Note 1: PCs with a small screen can fail to visualize the whole GUI; 
in this case, the user can add scroll bars with ```MODIStsp(scrollWindow=TRUE)```); 
Note 2: At the first execution of `MODIStsp`, a Welcome screen will appear, signaling
that `MODIStsp` is searching for a valid GDAL installation. Press "OK" and wait 
for GDAL to be found. If nothing happens for a long time (e.g., several minutes),
`MODIStsp` (and in particular the gdalUtils package on which it relies) is not
finding a valid GDAL installation in the more common locations. To solve the problem:
1. Ensure that GDAL is properly installed in your system
2. (On Windows) If it is installed, verify that GDAL is in your system PATH. and 
that the _GDAL\_DATA_ environment variable is correctly set (You can find simple instructions [HERE](http://gisforthought.com/setting-up-your-gdal-and-ogr-environmental-variables/))
3. If nothing works, signal it in the [issues](https://github.com/lbusett/MODIStsp/issues)
GitHub page of `MODIStsp` and we'll try to help!  </font size="2">

The  GUI allows selecting all processing options required for the creation of the
desired MODIS time series. The main available processing options are described in
detail in the following.

```{r GUIfig, echo=FALSE, fig.align="center", fig.heigth=7.5, message=FALSE, warning=FALSE}
  library(png)
  library(grid)
  library(knitr)
  img <- readPNG("MAIN_GUI.png")
  grid.raster(img)
```

## **Selecting Processing Parameters**

### **_MODIS Product, Platform and Layers:_**

The top-most menus allow to specify details of the desired output time series:

1. **"Category"** and **"Product"**: Selects the MODIS product of interest;
2. **MODIS platform(s)**: Selects if only TERRA, only AQUA or Both MODIS platforms 
should be considered for download and creation of the time series;
3. **version**: Selects whether processing version 5 or 6 (when available) of MODIS
products has to be processed

After selecting the product and version, clicking the **"Change Selection"** 
button opens the *_Select Processing Layers_* GUI panel, from which the user **must** 
select which MODIS original layers and/or derived Quality Indexes (QI) and Spectral
Indexes (SI) layers should be processed: 

```{r proc_layers, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE, paged.print=FALSE}
  library(png)
  library(grid)
  img <- readPNG('SELECT_GUI.png')
  grid.raster(img)
```

- The left-hand frame allows to select which _original MODIS layers_ 
should be processed.
- The central frame allows to select which _Quality Indicators should be extracted_ 
from the original MODIS Quality Assurance layers. 
- For MODIS products containing surface reflectance data, the right-hand frame 
allows to select which additional _Spectral Indexes should be computed_. The 
following commonly used Spectral Indexes are available for computation by default:

<center>**Table II: List of default Spectral Indexes available in MODIStsp**</center>

| Index Acronym   | Index name and reference                     |
|-----------------|----------------------------------------------|
| NDVI            | Normalized Difference Vegetation Index [@Rouse1973]         |
| EVI             | Enhanced Vegetation Index [@Huete2002]                      | 
| SR              | Simple Ratio [@Tucker1979]                                  |
| NDFI            | Normalized Difference Flood Index [@Boschetti2014]         |
| NDII7 (NDWI)    | Normalized Difference Infrared Index - Band 7 [@HUNTJR1989] |
| SAVI            | Soil Adjusted Vegetation Index  [@Huete1988]                |
| NDSI            | Normalized Difference Snow Index [@Hall2002]               |
| NDII6           | Normalized Difference Infrared Index - band 6 [@HUNTJR1989]|
| GNDVI           | Green Normalized Difference Vegetation Index [@Gitelson1998]|
| RGRI            | Red Green Ratio Index [@Gamon1999]                        |
| GRVI            | Green-Red ratio Vegetation Index [@Tucker1979]             |


You can however **specify other SIs to be computed without modifying MODIStsp source code** 
by clicking on the _**"Add Custom Index"**_ button, which allow to provide info related 
to the new desired SI using a simple GUI interface. 
```{r indexfig, echo=FALSE, fig.height=2.3, fig.width=6, message=FALSE, warning=FALSE, fig.align="center"}
  library(png)
  library(grid)
  img <- readPNG('NEWIND_GUI.png')
  grid.raster(img)
```

Provided information (e.g., correct band-names, computable formula, etc...) is 
automatically checked upon clicking "Set New Index". On success, the new index 
is added in the list of available ones for all products allowing its computation. 
Clicking "Done !" returns to the main.

__NOTE__ All custom defined indexes can be removed by using the `MODIStsp_resetindexes()` 
function

### **_Download Method:_**

Select the method to be used for download. Available choices are: 

1.  **http**: download through ftp from NASA lpdaac http archive (http://e4ftl01.cr.usgs.gov). 
This requires providing a user name and password, which can be obtained by registering 
an account at the address [https://urs.earthdata.nasa.gov/profile](https://urs.earthdata.nasa.gov/profile);

3.  **offline**: this option allows to process/reprocess HDF files already available 
on the user's PC without downloading from NASA -- useful if the user already has an
archive of HDF images, or to reprocess data already downloaded via MODIStsp to create
time series for an additional layer (_It is fundamental that the HDFs are those 
directly downloaded from NASA servers_ ! (See [here](faq.html#working-with-already-downloaded-hdf-files) 
for additional details). 

Checking the **use_aria2c** option allows to accelerate the download from NASA
archives. This requires however that that the "aria2c" software is installed in 
your system. To download and install it, see: [https://aria2.github.io/](aria2.github.io/)

<font size="2"> _NOTE: The best performances are usually achieved using *http*, 
though that may vary depending on network infrastructure._ </font size="2">

### **_Processing Period:_**

Specify the starting and ending dates to be considered for the creation of the 
time in the series corresponding fields. Dates **must** be provided in the _yyyy--mm--dd_ 
format (e.g., 2015-01-31)

The **Period** drop-down menu allows to choose between two options:

1.  **full**: all available images between the starting and ending dates are 
downloaded and processed;

2.  **seasonal**: data is downloaded only for one part of the year, but for 
multiple years. For example, if the starting date is 2005-03-01 and the ending is
2010-06-01, only the images of March, April and May for the years between 2005
and 2010 will be downloaded. This allows to easily process data concerning a 
particular season of interest.

### **_Spatial Extent:_**

Allows to define the area of interest for the processing. Two main options are possible: 

1.  **Select MODIS Tiles**: specify which MODIS tiles need to be processed either by: 

    a. Using the "Start" and "End" horizontal and vertical sliders in the
    _Required MODIS Tiles_ frame.  
    b. pressing the __"Select on Map"__ button. A map will open in a browser window,
    allowing interactive selection of the required tiles
    
During processing, data from the different tiles is mosaiced, and a single file 
covering the total area is produced for each acquisition date

2.  **Define Custom Area**: specify a custom spatial extent for the desired outputs 
either by: 
    
    a. Manually inserting the coordinates of the Upper Left and Lower Right corners 
    of the area of interest in the __Bounding Box__ frame. _Coordinates of the corners
    must be provided in the coordinate system of the selected  output projection_; 
    
    b. pressing the __"Load Extent from a Spatial File" and selecting a raster or 
    vector spatial file__. In this case, the bounding box of the selected file is
    retrieved, converted in the selected output projection, and shown in the
    "Bounding Box" frame. Required input MODIS tiles are also automatically retrieved 
    from the output extent, and the tiles selection sliders modified accordingly.
    
    c. pressing the __"Select on Map"__ button. A map will open in a browser window,
    allowing interactive selection of the spatial extent using the tools on the left.
    
    
**Note: ** pressing the "show current extent" will open a browser window highlighting
the currently selected spatial extent.
 
### _Reprojection and Resize_

Specify the options to be used for reprojecting and resizing the MODIS images. 

- **"Output Projection"**: select either the Native MODIS projection (Default) 
or specify a user-defined one. To specify a user selected projection, select 
"User Defined" and then insert a valid "Proj4" string in the pop-up window. 
Validity of the Proj4 string is automatically checked, and error messages issued 
if the check fails; 

- **"Output Resolution"**, **"Pixel Size"** and **"Reprojection Method"**: specify 
whether output images should inherit their spatial resolution from the original
MODIS files, or be resampled to a user-defined resolution. In the latter case, 
output spatial resolution must be specified in the measure units of the selected
output projection. Resampling method can  be chosen among "Nearest Neighbour" and 
"Mode" (Useful for down-sampling purposes). Other resampling methods (e.g., 
bilinear, cubic) are not currently supported since i) they cannot be used for
resampling of categorical variables such as the QA and QI layers, and ii) using 
them on continuous variable (e.g., reflectance, VI values) without performing an 
a-priori data cleaning would risk to contaminate the values of high-quality
observations with those of low-quality ones.

### _Output Options_

Several processing options can be set using check-boxes:

- **Output Files Format**: Two of the most commonly formats used in remote
sensing applications are available at the moment: ENVI binary and GeoTiff. If
GeoTiff is selected, the type of file compression can be also specified among
"None", "PACKBITS", "LZW" and "DEFLATE".

- **Save Time Series as**: Specify if virtual multitemporal files  should be 
created. These virtual files allow access to the entire time series of images as 
a single file without the need of creating large multitemporal raster images.
Available virtual files formats are "R" rasterStacks, ENVI meta-files and GDAL 
"vrt" files. In particular, "R" rasterStacks may be useful in order to easily 
access the preprocessed MODIS data within "R" scripts 
(see also http://lbusett.github.io/MODIStsp/articles/output.html).


- **Modify No Data**: Specify if NoData values of MODIS layers should be kept
at their original values, or changed to those specified within the "MODIStsp\_Products\_Opts"
XML file. By selecting "Yes" in the "Change Original NoData values" check-box, 
NoData of outputs are set to the largest integer value possible for the data type 
of the processed layer (e.g., for 8-bit  unsigned integer layers, NoData is set 
always to 255, for 16-bit signed  integer layers to 32767, and  for 16-bit unsigned
integer layers to 65535). Information about the new NoData values is stored both 
in the output rasters, and in the XML files associated with them. __NOTE:__ Some 
MODIS layers have multiple NoData (a.k.a. _fill_) values. if _Modify No Data_ is 
set to "Yes", `MODIStsp` will convert all _fill_ values to a common output NoData
value!

- **Apply Scale/Offset**: Specify if scale and offset values of the different 
MODIS layers should be applied. If selected, outputs are appropriately rescaled
on the fly, and saved in the true "measure units" of the selected parameter (e.g., 
spectral indexes are saved as floating point values; Land Surface Temperature is 
saved in degrees Kelvin, etc.). 

### __Main MODIStsp Output Folder__

Select the main folder where the pre-processed time series data will be stored. 
All `MODIStsp` outputs **will be placed in specific sub-folders of this main folder** (see http://lbusett.github.io/MODIStsp/articles/output.html for details on `MODIStsp` 
naming conventions)-.

The **"Reprocess Existing Data"** check-box allows to decide if images already 
available should be reprocessed if a new run of MODIStsp is launched with the same 
output folder. If set to "No", MODIStsp skips dates for which output files following
the MODIStsp naming conventions are already present in the output folder. This
allows to incrementally extend MODIS time series without reprocessing already available
dates. 

### _Folder for permanent storage of original MODIS HDF images_

Select the folder where downloaded **original MODIS HDF files** downloaded from 
NASA servers will be stored. 

The **"delete original HDF files"** check-box allows also to decide if the
downloaded images should be deleted from the file system at the end of the 
processing. To avoid accidental file deletion, this is always set to "No" by default, 
and a warning is issued before execution whenever the selection is changed to "Yes".

## **Saving and Loading Processing Options**

Specified processing parameters can be saved to a JSON file for later use by clicking
on the _**Save Options**_ button.

Previously saved options can be restored clicking on the _**Load Options**_ button
and navigating to the previously saved JSON file.

<font size="2"> (Note that at launch, **`MODIStsp` _always reloads automatically
the processing options used for its last successful run_** </font size="2">.

## **Starting the processing**

Once you are happy with your choices, click on **Start Processing**. `MODIStsp` 
will start accessing NASA servers to download and process the MODIS data corresponding 
to your choices.

For each date of the specified time period, `MODIStp` downloads and preprocesses
all hdf images required to cover the desired spatial extent. Informative messages
concerning the status of the processing are provided on the console, as well as on
a self-updating progress window. 

The processed time series are saved in specific subfolders of the main selected
output folder, as explained in detail [HERE](articles/output.html).

# Non-Interactive Execution from within R

## Specifying a saved "Options file"
`MODIStsp` can be launched in non-interactive mode within an `R` session by setting 
the optional `GUI` parameter to FALSE, and the `Options_File` parameter to the path 
of a previously saved JSON Options file. 
This allows to exploit `MODIStsp` functionalities within generic "R" processing 
scripts: 

```{r eval=FALSE, message=FALSE, warning=FALSE}
library(MODIStsp) 

# **NOTE** Output files of examples are saved to file.path(tempdir(), "MODIStsp").
# You can run the examples with `gui = TRUE` to set a different output folder!

# --> Specify the path to a valid options file saved in advance from MODIStsp GUI 
# Here we use a test json file saved in MODIStsp installation folder which
# downloads and processed 3 MOD13A2 images over the Como Lake (Lombardy, Italy)
# and retrieves NDVI and EVI data, plus the Usefulness Index Quality Indicator.

options_file <- system.file("testdata/test_MOD13A2.json", package = "MODIStsp")

# --> Launch the processing
MODIStsp(gui = FALSE, options_file = options_file, verbose = FALSE)

# Outputs are in this case in subfolder "MODIStsp/VI_16Days_1Km_v6" of `base::tempdir()`: 

out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6") 
list.files(out_fold)
#> [1] "EVI"         "NDVI"        "QA_usef"     "Time_Series"

list.files(file.path(out_fold ,"EVI"))
#> [1] "MOD13A2_EVI_2016_161.tif" "MOD13A2_EVI_2016_177.tif"
```

## Looping over different Options files

If you need to process different MODIS products, you can prepare beforehand different
MODIStsp options files by using the GUI, and then loop over them like this:

```{r eval=FALSE, message=FALSE, warning=FALSE}

opts_files <- c(system.file("testdata/test_MOD13A2.json", package = "MODIStsp"), 
                system.file("testdata/test_MOD10A2.json", package = "MODIStsp"))

for (opts_file in opts_files) {
  MODIStsp(gui = FALSE, options_file = opts_file)
}

# MOD13A2 ouputs
out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6") 
list.files(out_fold)
#> [1] "EVI"         "NDVI"        "QA_usef"     "Time_Series"
list.files(file.path(out_fold ,"EVI"))
#> [1] "MOD13A2_EVI_2016_161.tif" "MOD13A2_EVI_2016_177.tif"

# MOD10A2 ouputs
out_fold <- file.path(tempdir(), "MODIStsp/Surf_Temp_8Days_1Km_v6") 
list.files(out_fold)
#> [1] "Days_Clear"    "LST_Day_1km"   "LST_Night_1km" "Time_Series"  

list.files(file.path(out_fold ,"LST_Night_1km"))
#> [1] "MOD11A2_LST_Night_1km_2016_153.tif" "MOD11A2_LST_Night_1km_2016_161.tif"
#> [3] "MOD11A2_LST_Night_1km_2016_169.tif" "MOD11A2_LST_Night_1km_2016_177.tif"
```


## Looping on different spatial extents

Specifying also the `spatial_file_path_` parameter overrides for example the output 
extent of the selected Options File. This allows to perform the same preprocessing 
on different extents using a single Options File. For example:

```{r eval=FALSE, message=FALSE, warning=FALSE}

# Run the tool using the settings previously saved in a specific option file
# and specifying the extent from a spatial file allows to re-use the same
# processing settings to perform download and reprocessing on a different area
library(MODIStsp) 
options_file <- system.file("testdata/test_MOD13A2.json", package = "MODIStsp")
spatial_file <- system.file("testdata/lakeshapes/garda_lake.shp", package = "MODIStsp")
MODIStsp(gui = FALSE, options_file = options_file,
         spatial_file_path = spatial_file, verbose = FALSE)

# --> Create a character array containing a list of shapefiles (or other spatial files)
extent_list = list.files(system.file("testdata/lakeshapes/", package = "MODIStsp"), 
                         full.names = TRUE, "\\.shp$")  
extent_list

#> [1] "D:/Documents/Source/git/MODIStsp/inst/testdata/lakeshapes/garda_lake.shp"
#> [2] "D:/Documents/Source/git/MODIStsp/inst/testdata/lakeshapes/iseo_lake.shp"

# --> Loop on the list of spatial files and run MODIStsp using each of them to 
# automatically define the output extent (A separate output folder is created for
# each input spatial file).

for (single_shape in extent_list) {
  MODIStsp(gui = FALSE, options_file = options_file, 
           spatial_file_path = single_shape )
}

# output files are placed in separate folders. For example here you can see how to 
# access a list of produced files and plot one of them: 
outfiles_garda <- list.files(
  file.path(tempdir(), "MODIStsp/garda_lake/VI_16Days_1Km_v6/EVI"),
  full.names = TRUE)
library(raster)       
plot(raster(outfiles_garda[1]))

outfiles_iseo <- list.files(
  file.path(tempdir(), "MODIStsp/iseo_lake/VI_16Days_1Km_v6/EVI"),
  full.names = TRUE)
plot(raster(outfiles_iseo[1]))

```

# Stand-alone execution and scheduled processing  

## Stand-alone execution 

- MODIStsp can be also executed as a "stand-alone" application(i.e., without having
to open R/RStudio); to do this, from R launch the function `MODIStsp_install_launcher()`.

In a Linux operating system this function creates a desktop entry (accessible from
the menu in the sections "Science" and "Geography") and a symbolic link in a known path (default: /usr/bin/MODIStsp).
In Windows, a link in the Start Menu and optionally a desktop shortcut are created.
See `?install_MODIStsp_launcher` for details and path customization.

Double-clicking those files or launching them from a shell without parameters
will launch `MODIStsp` in interactive mode. Non-interactive mode is triggered
by adding the "-g"  argument to the call, and specifying the path to a valid 
Options File as "-s" argument:

  - Linux: `MODIStsp -g -s "/yourpath/youroptions.json"`
  (see `MODIStsp -h` for details).
  
  - Windows:`your_r_library\MODIStsp\ExtData\Launcher\MODIStsp.bat -g -s "yourpath/youroptions.json"`
  (see `C:\Users\you\Desktop\MODIStsp -h` for details).

If you do not want to install any link, launchers can be found in the subdirectory
"MODIStsp/ExtData/Launcher" of your library path.

## Scheduled Processing

Stand-alone non-interactive execution can be used to periodically and automatically 
update the time series of a selected product over a given study area. To do that, 
you should simply:

1.	Open the `MODIStsp` GUI, define the parameters of the processing specifying 
a date in the future as the "Ending Date" and save the processing options. Then
quit the program.
 
2. Schedule non-interactive execution of the launcher installed as seen before
(or located in the subdirectory "MODIStsp/ExtData/Launcher" of your library path)
as windows scheduled task (or linux "cron" job) according to a specified time 
schedule, specifying the path of a previously saved Options file as additional
argument:

#### On Linux

- edit your crontab by opening a terminal and type:

```bash
  crontab -e
```
 
- add an entry for the launcher. For example, if you have installed it in 
/usr/bin and you want to run the tool every day at 23.00, add the following row:
        
```bash
  0 23 * * * /bin/bash /usr/bin/MODIStsp -g -s "/yourpath/youroptions.json"
```
      
#### On Windows

- create a Task following [these instructions](https://technet.microsoft.com/en-us/library/cc748993.aspx); add the path of the MODIStsp.bat launcher as Action (point 6), and specifying  `-g -s "X:/yourpath/youroptions.json"` as argument.

# Outputs Format and Naming Conventions

## Single-band outputs

Output raster files are saved in specific subfolders of the main output folder. 
In particular, **a separate subfolder** is created for each processed original 
MODIS layer, Quality Indicator or Spectral Index. Each subfolder contains one image
for each processed date, created according to the following naming conventions: 

```
myoutfolder/"Layer"/"ProdCode"_"Layer"_"YYYY"_"DOY"."ext"
```

<font size="2"> _(e.g., myoutfolder/NDVI/MOD13Q1\_NDVI\_2000\_065.dat)_ </font size="2">

, where: 

  - **_Layer_** is a short name describing the dataset (e.g., b1_Red, NDII, UI);
  - **_ProdCode_** is the code name of the MODIS product from which the image was
  derived (e.g., MOD13Q1);
  - **_YYYY_** and **_DOY_** correspond to the year and DOY (Day of the Year) of 
  acquisition of the original MODIS image;
  - **_ext_** is the file extension (.tif for GTiff outputs, or .dat for ENVI outputs). 

____________________________________________________________________________________

## Virtual multi-band outputs

ENVI and/or GDAL virtual time series files and _RasterStack_ RData objects are 
instead stored **in the "Time\_Series" subfolder** if required.

Naming convention for these files is as follow:

```
myoutfolder/Time_Series/"vrt_type"/"Sensor"/"Layer"/"ProdCode"_"Layer"_"StartDOY"_"StartYear_"EndDOY"_"EndYear_"suffix".ext" 
```
<font size="2"> _(e.g., myoutfolder/Time_Series/RData/Terra/NDVI/MOD13Q1_MYD13Q1_NDVI_49_2000_353_2015_RData.RData)_ </font size="2"> 
             
, where: 

  - **_vrt_type_** indicates the type of virtual file ("RData", "GDAL" or "ENVI_META");
  - **_Sensor_** indicates to which MODIS sensor the time series belongs ("Terra",
  "Aqua", "Mixed" or "Combined" (for MCD* products));
  - **_Layer_** is a short name describing the dataset (e.g., b1_Red, NDII, UI);
  - **_ProdCode_** is the code name of the MODIS product from which the image was 
  derived (e.g., MOD13Q1);
  - **_StartDOY_**, **_StartYear_**, **_EndDOY_** and **_EndYear_** indicate the 
  temporal extent of the time serie created; 
  - **_suffix_** indicates the type of virtual file (ENVI, GDAL or RData);
  - **_ext_** is the file extension (".vrt" for gdal virtual files, "META" for 
  ENVI meta files or "RData" for `R` raster stacks). 


# Accessing the processed time series from R

Preprocessed MODIS data can be retrieved within `R`  either by accessing the single-date raster files, or by loading the saved _RasterStack_ objects. 

Any single-date image can be accessed by simply opening it with a `raster` command: 

```{R eval=FALSE, tidy=TRUE, highlight=TRUE}
library(raster)
modistsp_file <- "/my_outfolder/EVI/MOD13Q1_2005_137_EVI.tif"
my_raster     <- raster(modistsp_file)
```

`rasterStack` time series containing all the processed data for a given parameter (saved in the "Time Series/RData" subfolder - see [here](output.html) for details) can be opened by: 

```{r eval=FALSE}
in_virtual_file <- "/my_outfolder/Time_Series/RData/Terra/EVI/MOD13Q1_MYD13Q1_NDVI_49_2000_353_2015_RData.RData" 
indata          <- get(load(in_virtual_file))
```

This second option allows accessing the complete data stack and analyzing it using the functionalities for raster/raster time series analysis, extraction and plotting provided for example by the ```raster``` or   ```rasterVis``` packages.


## Extracting Time Series Data on Areas of Interest

`MODIStsp` provides an efficient function (`MODIStsp\_extract`) for extracting time series data at specific locations. The function takes as input a _RasterStack_ virtual object created by `MODIStsp` (see above), the starting and ending dates for the extraction and a standard _Sp*_ object (or an ESRI shapefile name) specifying the locations (points, lines or polygons) of interest, and provides as output a `xts` object or `data.frame` containing time series data for those locations. 

If the input is of class _SpatialPoints_, the output object contains one column for each point specified, and one row for each date. If it is of class _SpatialPolygons_ (or _SpatialLines_), it contains one column for each polygon (or each line), with values obtained applying the function specified as the "FUN" argument (e.g., mean, standard deviation, etc.) on pixels belonging to the polygon (or touched by the line), and one row for each date. 

As an example the following code:

```{r, eval=FALSE}
  #Set the input paths to raster and shape file
  infile    <- 'myoutfolder/Time_Series/RData/Mixed/MOD13Q1_MYD13Q1_NDVI_49_2000_353_2015_RData.RData'  
  shpname   <- 'path_to_file/rois.shp'  
  #Set the start/end dates for extraction
  startdate <- as.Date("2010-01-01")  
  enddate   <- as.Date("2014-12-31")    
  #Load the RasterStack
  inrts     <- get(load(infile)) 
  # Compute average and St.dev
  dataavg   <- MODIStsp_extract(inrts, shpname, startdate, enddate, FUN = 'mean', na.rm = T)
  datasd    <- MODIStsp_extract (inrts, shpname, startdate, enddate, FUN = 'sd', na.rm = T)
  # Plot average time series for the polygons
  plot.xts(dataavg) 
```

  loads a _RasterStack_ object containing 8-days 250 m resolution time series for the 2000-2015 period and extracts time series of average and standard deviation values over the different polygons of a user's selected shapefile on the 2010-2014 period.
  
# Problems and Issues

Solutions to some common **installation and processing problems** can be found in MODIStsp FAQ: 

http://lbusett.github.io/MODIStsp/articles/faq.html

- Please **report any issues** you may encounter in our issues page on GitHub:

https://github.com/lbusett/MODIStsp/issues

  
# Citation
  
To cite MODIStsp please use:

L. Busetto, L. Ranghetti (2016) MODIStsp: An R package for automatic preprocessing of MODIS
  Land Products time series, Computers & Geosciences, Volume 97, Pages
  40-48, ISSN 0098-3004, http://dx.doi.org/10.1016/j.cageo.2016.08.020, URL: https://github.com/lbusett/MODIStsp. 
  

<a name="gdal_inst"/>

# Installing R and GDAL

## Installing R

### Windows

Download and install the latest version of R which can be found [here](https://cran.r-project.org/bin/windows/base).

### Linux

Please refer to the documentation which can be found [here](https://cran.r-project.org/bin/linux), opening the directory relative to your Linux distribution. The documentation provides instruction to add CRAN repositories and to install the latest R version.
With Ubuntu 15.10 Wily (and newer) this step is not mandatory (although recommended), since packaged version of R is >= 3.2.1 (although not the latest); in this case, user can install R by simply typing in a terminal
```bash
sudo apt-get install r-base
```

## Installing GDAL >= 1.11.1

### Windows

The easiest way to install GDAL on Windows is from the [OSGeo4W Website](https://trac.osgeo.org/osgeo4w/)

1. Open the [OSGeo4W Website](https://trac.osgeo.org/osgeo4w/)
2. In the **Quick Start for OSGeo4W Users** section, select the download of 32bit or 64bit of OSGeo4W network installer
3. Run the installer
  + _Easiest Option_: 
    +  Select **Express Desktop Install**, then proceed with the installation. This will install GDAL and also other useful Spatial Processing software like [QGIS](http://www.qgis.org/) and [GRASS GIS](https://grass.osgeo.org/)
  + _Advanced Option_: 
    +  Select **Advanced Install**, then click on "Next" a few times until you reach the "Select Packages" screen. 
    +  Click on "Commandline_Utilities_", and on the list look for "_gdal: The GDAL/OGR library..." entry
    +  Click on "Skip": the word "skip" will be replaced by the current GDAL version number
    +  Click on "Next" a few times to install GDAL
    
### Debian and Ubuntu-based systems

1. Ensure that your repositories contain a version of ```gdal-bin``` >= 1.11.1. 
    In particular, official repositories of Ubuntu 15.04 Vivid (or older) and Debian Jessie (or older) provide older versions of GDAL, so it is necessary to add UbuntuGIS-unstable repository before installing. To do this, follow instructions [here](https://launchpad.net/~ubuntugis/+archive/ubuntu/ubuntugis-unstable)). 
    With Ubuntu 15.10 Wily (and newer) this step is not mandatory, although recommended in order to have updated version of GDAL installed.
    
2. To install GDAL, open a terminal and type  
    ```bash
    sudo apt-get install gdal-bin
    ```
    
### ArchLinux

GDAL is maintained updated to the latest version as binary package within the community repository; although that, the support for HDF4 format is not included. 
To bypass this problem, ArchLinux users can install ```gdal-hdf4``` package from AUR (see [here](https://wiki.archlinux.org/index.php/Arch_User_Repository#Installing_packages) or [here](https://archlinux.fr/yaourt-en) for the package installation from AUR). 
This package is updated manually after each release of ```gdal``` on the community repository, so a temporal shift between a new ```gdal``` release and the update of ```gdal-hdf4``` could happen.
If you want to manually add the support for HDF4 in case ```gdal-hdf4``` is out-of-date, you can do it following [these instructions](https://notehub.org/fctdn).

### Other Linux systems

Install the packaged binary of GDAL included in your specific distribution; if the version is older than 1.11.1, or if the support for HDF4 format is not included, you can manually install the HDF4 library and compile the source code by adding the parameter ```--with-hdf4``` to the ```configure``` instruction).

#### MacOs

1. Check if gdal is already installed and has hdf4 support. In a terminal, type: 

```
~$ gdal-config --formats 
```
- if gdal is installed, the command will list which drivers are installed. If the 
list includes "hdf4" you should be good to go. 

- if gdal is not yet installed or hdf4 is not already supported, you can install/reinstall 
it following these [notes](http://r-sig-geo.2731867.n2.nabble.com/OSX-rgdal-GDAL-drivers-and-HDF4-td7588400.html). In short: open a terminal and run:

```
~$ brew install hdf4 
~$ brew link --overwrite hdf4 
~$ brew install gdal --complete --enable-unsupported --with-hdf4 
```

then, check again to see if hdf4 is supported. The list should now include hdf4: 

```
~$ gdal-config --formats 
```

Good luck!

[^1^]: If you encounter problems installing the gWidgetsRGtk2 library, please signal it in the [issues](https://github.com/lbusett/MODIStsp/issues) GitHub page of MODIStsp and we'll try to help you!

[^2^]: At the first execution of _MODIStsp_, a Welcome screen will appear, signaling that MODIStsp is searching for a valid GDAL installation. Press "OK" and wait for GDAL to be found. If nothing happens for a long time (e.g., several minutes), MODIStsp (and in particular the gdalUtils package on which it relies) is not finding a valid GDAL installation in the more common locations. To solve the problem: i) ensure that GDAL is properly installed in your system; ii) (on Windows) if it is installed, verify that GDAL is in your system PATH. and that the _GDAL\_DATA_ environment variable is correctly set (You can find simple instructions [HERE](http://gisforthought.com/setting-up-your-gdal-and-ogr-environmental-variables/)); iii) if nothing works, signal it in the [issues](https://github.com/lbusett/MODIStsp/issues) GitHub page of MODIStsp and we'll try to help!

[^3^]: The lists of original MODIS layers, QIs and Sis available for the selected product are automatically retrieved from the "MODIStsp_Products_Opts" XML file distributed with the package in /ExtData subfolder.

# References
