---
title: "Non-Interactive Execution from within R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Launching MODIStsp using a saved "Options file"__

`MODIStsp` can be launched in non-interactive mode within an `R` session by setting 
the optional `GUI` parameter to FALSE, and the `Options_File` parameter to the path 
of a previously saved JSON Options file. 
This allows to exploit `MODIStsp` functionalities within generic "R" processing 
scripts: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(MODIStsp) 

# **NOTE** Output files of examples are saved to file.path(tempdir(), "MODIStsp").
# You can run the examples with `gui = TRUE` to set a different output folder!

# --> Specify the path to a valid options file saved in advance from MODIStsp GUI 
# Here we use a test json file saved in MODIStsp installation folder which
# downloads and processed 3 MOD13A2 images over the Como Lake (Lombardy, Italy)
# and retrieves NDVI and EVI data, plus the Usefulness Index Quality Indicator.

options_file <- system.file("testdata/test_MOD13A2.json", package = "MODIStsp")

# --> Launch the processing
MODIStsp(gui = FALSE, options_file = options_file, verbose = FALSE)

# Outputs are in this case in subfolder "MODIStsp/VI_16Days_1Km_v6" of 
# `base::tempdir()`: 

out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6") 
list.files(out_fold)
list.files(file.path(out_fold ,"EVI"))
```
________________________________________________________________________________

<br>

#<i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Looping over different Options files__

If you need to process different MODIS products, you can prepare beforehand different
MODIStsp options files by using the GUI, and then loop over them like this:

```{r echo=TRUE, message=FALSE, warning=FALSE}

opts_files <- c(system.file("testdata/test_MOD13A2.json", package = "MODIStsp"), 
                system.file("testdata/test_MOD10A2.json", package = "MODIStsp"))

for (opts_file in opts_files) {
  MODIStsp(gui = FALSE, options_file = opts_file)
}

# MOD13A2 ouputs
out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6") 
list.files(out_fold)
list.files(file.path(out_fold ,"EVI"))

# MOD10A2 ouputs
out_fold <- file.path(tempdir(), "MODIStsp/Surf_Temp_8Days_1Km_v6") 
list.files(out_fold)
list.files(file.path(out_fold ,"LST_Night_1km"))
```
________________________________________________________________________________

<br>

#<i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Looping over different spatial extents__

Specifying also the `spatial_file_path_` parameter overrides for example the output 
extent of the selected Options File. This allows to perform the same preprocessing 
on different extents using a single Options File. For example:

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Run the tool using the settings previously saved in a specific option file
# and specifying the extent from a spatial file allows to re-use the same
# processing settings to perform download and reprocessing on a different area
library(MODIStsp) 
options_file <- system.file("testdata/test_MOD13A2.json", package = "MODIStsp")
spatial_file <- system.file("testdata/lakeshapes/garda_lake.shp", package = "MODIStsp")
MODIStsp(gui = FALSE, options_file = options_file,
         spatial_file_path = spatial_file, verbose = FALSE)

# --> Create a character array containing a list of shapefiles (or other spatial files)
extent_list = list.files(system.file("testdata/lakeshapes/", package = "MODIStsp"), 
                         full.names = TRUE, "\\.shp$")  
extent_list

# --> Loop on the list of spatial files and run MODIStsp using each of them to 
# automatically define the output extent (A separate output folder is created for
# each input spatial file).

for (single_shape in extent_list) {
  MODIStsp(gui = FALSE, options_file = options_file, 
           spatial_file_path = single_shape )
}

# output files are places in separate folders: 
outfiles_garda <- list.files(
  file.path(tempdir(), "MODIStsp/garda_lake/VI_16Days_1Km_v6/EVI"),
  full.names = TRUE)
library(raster)       
plot(raster(outfiles_garda[1]))

outfiles_iseo <- list.files(
  file.path(tempdir(), "MODIStsp/iseo_lake/VI_16Days_1Km_v6/EVI"),
  full.names = TRUE)
plot(raster(outfiles_iseo[1]))

```
