---
title: "Non-Interactive Execution from within R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`MODIStsp` can be launched in non-interactive mode within an `R` session by setting the optional `GUI` parameter to FALSE, and either providing the desired processing argument in the call to the function, or providing a previously saved `opts_file` specifying the path to a JSON Options file previously saved through the GUI. 
This allows to exploit `MODIStsp` functionalities within generic "R" processing scripts.

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Specifying the processing parameters in the function call__

All processing parameters can be set in the call to `MODIStsp`. Mandatory 
parameters are `selprod` (specifying the MODIS product), (one of) `bandsel`, `quality_bandsel` or `indexes_bandsel`, (that specify the desired output layers), `out_folder` and `start_date` and `end_date`. 
`user` and `password` are also needed if `download_server` is not equal to `offline`.

The new function `MODIStsp_get_prodlayers` allows to easily retrieve the 
names of products and available layers based on product code, such as in: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(MODIStsp) 
MODIStsp_get_prodlayers("M*D13Q1")
```

The other parameters are set automatically to default values (see [HERE](reference/index.html) for details on the different available function arguments). 

For example, the following code processes layers __NDVI__ and __EVI__ and quality indicator __usefulness__ of product __M*D13Q1__, considering both Terra and Aqua platforms, for dates comprised between 2020-06-01 and 2020-06-15 and saves output to R tempdir: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(MODIStsp) 

# **NOTE** Output files of examples are saved to file.path by setting out_folder to $tempdir.

# --> See name and available layers for product M*D13Q1
# MODIStsp_get_prodlayers("M*D13A2")

# --> Launch the processing
MODIStsp(gui             = FALSE, 
         out_folder      = "$tempdir", 
         selprod         = "Vegetation_Indexes_16Days_1Km (M*D13A2)",
         bandsel         = c("EVI", "NDVI"), 
         quality_bandsel = "QA_usef", 
         indexes_bandsel = "SR", 
         user            = "mstp_test" ,
         password        = "MSTP_test_01",
         start_date      = "2020.06.01", 
         end_date        = "2020.06.15", 
         verbose         = FALSE)

# Outputs are in this case in subfolder "MODIStsp/VI_16Days_1Km_v6" of 
# `base::tempdir()`: 

out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6/") 
list.files(out_fold)
list.files(file.path(out_fold ,"EVI"))
list.files(file.path(out_fold ,"QA_usef"))
```

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Launching MODIStsp using a saved "Options file"__

Alternatively, you can run MODIStsp without opening the GUI by specifying a previously saved options file: 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(MODIStsp) 

# **NOTE** Output files of examples are saved to file.path(tempdir(), "MODIStsp").

# --> Specify the path to a valid options file saved in advance from MODIStsp GUI 
# Here we use a test json file saved in MODIStsp installation folder which
# downloads and processed 3 MOD13A2 images over the Como Lake (Lombardy, Italy)
# and retrieves NDVI and EVI data, plus the Usefulness Index Quality Indicator.

opts_file <- system.file("testdata/test_MOD13A2.json", package = "MODIStsp")

# --> Launch the processing
MODIStsp(gui = FALSE, opts_file = opts_file, verbose = FALSE)

# Outputs are in this case in subfolder "MODIStsp/VI_16Days_1Km_v6" of 
# `base::tempdir()`: 

out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6") 
list.files(out_fold)
list.files(file.path(out_fold ,"EVI"))
```

## _Looping over different Options files_

If you need to process different MODIS products, you can prepare beforehand different
MODIStsp options files by using the GUI, and then loop over them like this:

```{r echo=TRUE, message=FALSE, warning=FALSE}

opts_files <- c(system.file("testdata/test_MOD13A2.json", package = "MODIStsp"), 
                system.file("testdata/test_MOD10A2.json", package = "MODIStsp"))

for (opts_file in opts_files) {
  MODIStsp(gui       = FALSE, 
           opts_file = opts_file, 
           verbose   = FALSE)
}

# MOD13A2 ouputs
out_fold <- file.path(tempdir(), "MODIStsp/VI_16Days_1Km_v6") 
list.files(out_fold)
list.files(file.path(out_fold ,"NDVI"))

# MOD10A2 ouputs
out_fold <- file.path(tempdir(), "MODIStsp/Surf_Temp_8Days_1Km_v6") 
list.files(out_fold)
list.files(file.path(out_fold ,"LST_Night_1km"))
```

________________________________________________________________________________

<br>

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Specifying the processing parameters using a previously saved options file and overwriting some parameters__

Finally, it is possible to both specify a previously saved options file AND setting some parameters in the call to the function. This allows to __easily perform similar processings, by only updating the required arguments__, as in the examples below. 

## __Looping over different spatial extents__

Specifying the `spafile` parameter while setting the `spatmeth` parameter to "file" overrides for example the output extent of the selected Options File. This allows to perform the same preprocessing on different extents using a single Options File. For example:

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Run the tool using the settings previously saved in a specific option file
# and specifying the extent from a spatial file allows to re-use the same
# processing settings to perform download and reprocessing on a different area
library(MODIStsp) 
opts_file    <- system.file("testdata/test_MOD13A2.json", package = "MODIStsp")
spatial_file <- system.file("testdata/lakeshapes/garda_lake.shp", package = "MODIStsp")
MODIStsp(gui       = FALSE, 
         opts_file = opts_file,
         spatmeth  = "file", 
         spafile   = spatial_file, 
         verbose   = FALSE)

# --> Create a character array containing a list of shapefiles (or other spatial files)
extent_list <- list.files(system.file("testdata/lakeshapes/", package = "MODIStsp"), full.names = TRUE, "\\.shp$")  

extent_list

# --> Loop on the list of spatial files and run MODIStsp using each of them to 
# automatically define the output extent (A separate output folder is created for
# each input spatial file).

for (single_shape in extent_list) {
  MODIStsp(gui       = FALSE, 
           opts_file = opts_file, 
           spatmeth  = "file", 
           spafile   = single_shape, 
           verbose   = FALSE)
}

# output files are placed in separate folders: 

outfiles_garda <- list.files(
  file.path(tempdir(), "MODIStsp/garda_lake/VI_16Days_1Km_v6/EVI"),
  full.names = TRUE)
outfiles_garda

library(raster)       
plot(raster(outfiles_garda[1]))

outfiles_iseo <- list.files(
  file.path(tempdir(), "MODIStsp/iseo_lake/VI_16Days_1Km_v6/EVI"),
  full.names = TRUE)
outfiles_iseo

plot(raster(outfiles_iseo[1]))

```
