---
title: "Non-Interactive Execution from within R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# <i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Launching MODIStsp using a saved "Options file"__

`MODIStsp` can be launched in non-interactive mode within an `R` session by setting 
the optional `GUI` parameter to FALSE, and the `Options_File` parameter to the path 
of a previously saved JSON Options file. 
This allows to exploit `MODIStsp` functionalities within generic "R" processing scripts

```{r, eval=FALSE}
library(MODIStsp) 
# --> Specify the path to a valid options file saved in advance from MODIStsp GUI 
options_file <- "X:/yourpath/youroptions.json" 
  
# --> Launch the processing
MODIStsp(gui = FALSE, options_file = options_file)
```
________________________________________________________________________________

<br>

#<i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Looping over different Options files__

If you need to process different MODIS products, you can prepare beforehand different
MODIStsp options files by using the GUI, and then loop over them like this:

```{r, eval=FALSE}

opts_files <- c("X:/yourpath/youroptions_MOD13Q1.json",
                "X:/yourpath/youroptions_MOD10A1.json")

for (opts_file in opts_files) {
  MODIStsp(gui = FALSE, options_file = opts_file)
}
```
________________________________________________________________________________

<br>

#<i class="fa fa-arrow-circle-o-right" aria-hidden="true"></i> __Looping over different spatial extents__

Specifying also the `spatial_file_path_` parameter overrides for example the output 
extent of the selected Options File. This allows to perform the same preprocessing 
on different extents using a single Options File, by looping on an array of spatial 
files representing the desired output extents. For example:

```{r eval=FALSE}
library(MODIStsp) 
# --> Specify the path to a valid options file saved in advance from MODIStsp GUI 
options_file <- "X:/yourpath/youroptions.json" 

# --> Create a character array containing a list of shapefiles (or other spatial files)
extent_list <- list.files("X:/path/containing/some/shapefiles/", "\\.shp$")  

# --> Loop on the list of spatial files and run MODIStsp using each of them to 
# automatically define the output extent (A separate output folder is created for
# each input spatial file).

for (single_shape in extent_list) {
  MODIStsp(gui = FALSE, options_file = options_file, 
          spatial_file_path = single_shape )
}

```
